---
title: "R Notebook"
output: html_notebook
date: "21/03/2021"
Author: "Israel Usen"
---

#Data wrangling and exploration with multiple datasets
Notes: setting up my R environment by loading the 'tidyverse', 'palmer penguins' and lots of other packages for my exploration.

```{r}
#loading the dplyr package
library("dplyr")
```


```{r}
# This is a dataset that's pre-installed in R
data("ToothGrowth")
View(ToothGrowth)

```

```{r}
#only rows with a dose of 0.5 and is assigned to filtered_tg

filtered_tg <- filter(ToothGrowth,dose==0.5)
View(filtered_tg)
```

```{r}
#sorting filtered_tg in ascending order using the len column
arrange(filtered_tg, len)
```

```{r}
#using a nested function to run the same code above, this makes the code to run from the inside out

arrange(filter(ToothGrowth,dose==0.5),len)

```

```{r}
#using a pipe i.e. %>% 
filtered_toothgrowth <- ToothGrowth %>% 
  filter(dose==0.5) %>% 
  arrange(len)

View(filtered_toothgrowth) #presto!

```

```{r}
#to find the average length of tooth when the dose==0.5
filtered_toothgrowth <- ToothGrowth %>% 
  filter(dose==0.5) %>% 
  group_by(supp) %>% 
  na.omit() %>% #to remove na values
  summarize(mean_len=mean(len))

View(filtered_toothgrowth) #presto!!
```

```{r}
#The diamonds dataset
library("tidyverse")
library(ggplot2)
 
```

```{r}
#The diamonds dataset
data("diamonds")
View(diamonds)
```

```{r}
mutate(diamonds, carat_2=carat*100) #this code adds a new column but doesn't affect the original dataframe
```

```{r}
#creating plots for the diamonds dataset
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=color, fill=cut))+
  facet_wrap(~cut) #this split the plots by cuts
```

```{r}
#a basic bar chart
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut))
```

```{r}
#a basic bar chart with colour 
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut, color= cut))
```

```{r}
#a basic bar chart with colour fill mapped to cut
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut, fill= cut))
```

```{r}
#a basic bar chart with colour fill mapped to clarity
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut, fill= clarity))
  
```

```{r}
#Anscombe's quartet: four datasets that have nearly identical summary statistics
library('Tmisc')
data(quartet)
View(quartet)
```

```{r}
#to calculate mean and corelation 
quartet %>% 
  group_by(set) %>% 
  summarize(mean(x), sd(x), mean(y), sd(y), cor(x,y))
#the mean for x is 9 and the mean for y is 7.5, the correlation shows that this data is identical, which can be misleading
```

```{r}
#to plot graph and identify the difference
ggplot(quartet, aes(x,y)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE) +
  facet_wrap(~set) #this makes it plot for each set
#from this graph, it is clear that there are differences between each of the sets.
```

```{r}
#the datasauRus package create's plot with the Anscrombe data in different shapes
library(datasauRus)

```

```{r}

#creating a chart using the datasauRus package, using the Anscrombe dataset
ggplot(datasaurus_dozen, aes(x=x, y=y, color=dataset )) +
  geom_point() +
  theme_void() +
 # theme(legend.position = "none") +
  facet_wrap(~dataset, ncol=4)
# The trick to craeting these datasets is described in the paper by Justin Matejka and George Fitzmaurice "Same Stats, Different Graphs"

#using the datasauRus package makes it easier to spot the difference between all the datasets
```

```{r}
#Creating another plot using the datasauRus package
ds = data.frame()
for(j in 1:4){
  ds = rbind( ds, data.frame( number = j , x = anscombe[, j], y = anscombe[,j+4]))
}
ggplot(ds, aes(x, y) ) + geom_point( color = "red", fill="red", shape=17) + facet_wrap(~number)

#this shows plots for all four datasets
```


#using bias function it is present in a package called SimDesign
#the bias function determines the average the actual outcome is greater than the predicetd outcome, this is done by comparing both the actual outcome and the predicted outcome.

```{r}
install.packages("SimDesign")
library("SimDesign")
```

```{r}
#assuming this are data from a local weather channel
actual_temp<- c(68.3, 70, 72.4, 71, 67, 70) #assigning random values for sample test
predicted_temp<- c(67.9, 69, 71.5, 70, 67, 69)

bias(actual_temp, predicted_temp)
#the result is 0.71 which makes it close to zero, but it seems biased towards lower temperatures this isn't as accurate as it should be.
```

```{r}
#assuming this are data from a game store, that keeps record of how many games they sell on release day

actual_sales<- c(150, 203, 137, 247, 116, 287)
predicted_sales<- c(200, 300, 150, 250, 150,300)
bias(actual_sales, predicted_sales)

#this gave a -35, the predicted outcome is larger than the actual outcome, this means that they may have been ordering too much stocks for release days.
```



